---
title: "Stat242:Lab4"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
description: >
  Making Claims about a Population Based on Sample Data.
vignette: >
  %\VignetteIndexEntry{Stat242:Lab4}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  warning = FALSE
)
```

```{r setup, include = FALSE}
library(Stat242)
library(tidyverse)
library(infer)
```

## Making Claims about a Population Based on Sample Data

As we've seen in chapter 19, we can calculate confidence intervals and conduct hypothesis test to make claims about means in our population but based on our sample. In this lab, we'll build an interval and conduct a hypothesis test for a two-sample average and for many samples (chapter 22). 

For the first part, we'll be using the `piracy` data set included in  our `Stat242` package. It contains data for each member of congress and money they received from lobby groups for (money_pro) and against (money_con) anti-piracy legislation. 

## Means

For our means section Our test/interval will be to determine if there is a difference in money received by Democrats and Republicans, regardless of which lobby group the money came from.  

To clean our data, we'll need to replace missing values with 0 (rationale discussed in lab/the screencast), create a new variable for the sum of money received and filter our the few Independent party observations. 

```{r}
summary_data <- Stat242::piracy %>% 
  mutate(money_pro = replace_na(money_pro, 0), 
         money_con = replace_na(money_con,0),
         money_total = money_pro+money_con) %>% 
  select(-c(money_pro, money_con)) %>%  
  filter(party!= "I")
```

Now we're ready to run our test.

```{r}
t_test(summary_data, #choose the data frame
       formula = money_total~party, #define the relationship/variables for testing
       order = c("D","R"), #choose an order for the mean difference to be calculated
       mu = 0, #define the null hypothesis
       alternative = "two.sided", #define the alternative hypothesis
       var.equal = TRUE, #assume the populations have equal variance
       conf_int = TRUE, #will calculate a confidence interval for the mean difference as part of the test
       conf_level = 0.99) #defines the confidence/significance level
```

From the output, we can conclude there is a difference. Based on the interval is appears that Democrats received more money in relation to the anti-piracy legislation.

## ANOVA

The analysis of variance - anova - test is equivalent to the chi-squared test but for numeric data instead of categorical data; It will compare the means of three or more samples. 

We will be going back to the diamonds data set and running an anova test to see if there is a difference in the average price of diamonds based on clarity. We will nest the test inside a `summary()` function to get the test and print out of results in one line. 


```{r}
summary(aov(price ~ clarity, data = diamonds))
```

The important parts of this print out are the $F$ value and $Pr(>F)$. The $F$ value is similar to a $z$ or $t$ score. The $Pr(>F)$ is a p-value. 

Using the p-value method, $<2e-16$ means the p-value is less than $0.0000000000000002$. Using any of the typical values for alpha, we would reject our null hypothesis and can say there is a difference in the average price of diamonds based on their clarity rating.

Using the classic method, we need to calculate a critical F value. 

```{r}
qf(0.95, #1-alpha
   df1 = 7,
   df2 = 53932)
```

This means our critical region starts at approximately $F = 2$, and our test statistic is $F = 215$. We are well into the critical region and again conclude that there is a difference in the average price of diamonds based on their clarity. 