---
title: "Stat231:Lab2"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
description: >
  For this lab, we will be looking at calculating summary statistics, such as a five number summary, as well as boxplots.
vignette: >
  %\VignetteIndexEntry{Stat231:Lab2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE
)
```

```{r setup}
library(Stat231)
library(tidyverse)
library(openintro)
```

We will be using the diamonds dataset from the *openintro* package. To view the data either run the code chunk below or type the command into your Console. 

```{r, eval = FALSE}
View(diamonds)
```

## Five Number Summary

As you can see, this is a large data set with information on 53,940 individual diamonds! First lets calculate some summary statistics.

```{r}
summary(diamonds$price)
```
The *summary()* function outputs the five number summary, plus the mean, of a numeric variable.

Let's see what happens when we use the *summary()* function on a categorical variable. 

```{r}
summary(diamonds$color)
```
With a categorical variable, *summary()* outputs a frequency distribution!

With such a large data set, it might be useful to have some summary statistics for specific subsets. For example, we might expect diamonds with different clarity to have different prices. The *tidyverse* allows us to group our dataset by a specific variable before calculating summary values. For this, we'll use the *fivenum()* function. For example:

```{r}
fivenum(diamonds$price)[1]
```
Will give the minimum value (1st number of the summary) of the variable. 

```{r}
diamonds_summary <- diamonds %>% #This will create a new data set called diamonds_summary by taking the diamonds dataset "and then"...
  group_by(clarity) %>% 
  summarize(min = fivenum(price)[1],
            q1 = fivenum(price)[2],
            med = fivenum(price)[3],
            q3 = fivenum(price)[4],
            max = fivenum(price)[5]) 
```

## Boxplots
Let's create a few boxplots.

```{r}
ggplot(data = diamonds, 
       aes(y = price))+
  geom_boxplot()
```

What if we wanted the boxplot for the grouped data?

```{r}
ggplot(data = diamonds,
       aes(x = clarity, 
           y = price))+
  geom_boxplot()