---
title: "Stat242:Lab2"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
description: >
  Being able to analyze data is a good quality. Being able to present it in a clear, concise format is even better. First, we will focus expanding the graphs we made in lab 1 to be more clear, concise and have a bit more style. We'll finish learning how to create unique/custom styling.
vignette: >
  %\VignetteIndexEntry{Stat242:Lab2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE
)
```

```{r setup, include = FALSE}
library(stat242)
library(tidyverse)
library(openintro)
library(ggthemes) 
```

## Quality Data Visualization

Being able to analyze data is a good quality. Being able to present it in a clear, concise format is even better. In part I of our data viz labs, we will focus expanding the graphs we made in lab 1 to be more clear, concise and have a bit more style. 

First, lets draw our original histogram. 

```{r}
ggplot(data = gpa_study_hours, 
       aes(x = gpa))+
  geom_histogram()
```

Recall, we did make a few adjustments to this in lab 1; we changed the color of the border of the bars as well as their width. 

```{r}
ggplot(data = gpa_study_hours, 
       aes(x = gpa))+
  geom_histogram(color = "White", binwidth = 0.25)
```

It would definitely be helpful to add a title and clean up the labels a bit. Using the 'Help' tab in the Directory pane, look up our data set, `gpa_study_hours`, and read the description. We'll use this to write our labels. 

```{r}
ggplot(data = gpa_study_hours, 
       aes(x = gpa))+
  geom_histogram(color = "White", binwidth = 0.25)+
  labs( #the labs() function allows us to add labels to our graph
    title = "Distribution of Grade Point Average",
    subtitle = "For 193 Introductory Statistics Students",
    x = "GPA",
    y = "Number of Students",
    caption = "Source: Private US University, 2012"
  )
```

For our next step, we'll add a little flair. We'll use some styles or "themes" from the `ggthemes` package. Use the 'Packages' tab in the Directory pane to look up `ggthemes` and take a look at some of our options. Note: In part II of the lab we will create more custom styles. You can think of part I (this lab) as using a cookie recipe someone else made. We'll eat the cookie, see if we like it, and change what we don't in the next lab.

```{r}
ggplot(data = gpa_study_hours, 
       aes(x = gpa))+
  geom_histogram(color = "White", binwidth = 0.25)+
  labs(
    title = "Distribution of Grade Point Average",
    subtitle = "For 193 Introductory Statistics Students",
    x = "GPA",
    y = "Number of Students",
    caption = "Source: Private US University, 2012"
  )+
  theme_solarized()
```

I like this theme, but the gray bars seem out of place. I'll change the recipe a bit. 

```{r}
ggplot(data = gpa_study_hours, 
       aes(x = gpa))+
  geom_histogram(fill = "lightgoldenrod2",
                 color = "White", 
                 binwidth = 0.25)+
  labs(
    title = "Distribution of Grade Point Average",
    subtitle = "For 193 Introductory Statistics Students",
    x = "GPA",
    y = "Number of Students",
    caption = "Source: Private US University, 2012"
  )+
  theme_solarized()
```

That looks much better. To change the color of the bars, I added the `fill` argument to the geom_histogram() function. Here is a list of [colors](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf).

## Custom Data Visualization

For this part of lab, we will reset to the plot above that just had labels. Then instead of using a prebuilt theme from the `ggthemes` package, we will be creating our own unique visualizations.  

To customize a ggplot image, we'll add a `theme()` function to the end and change specific parts of the graph. Here is a [full list](https://ggplot2.tidyverse.org/reference/theme.html) of what you can change with the `theme()` function. It is a substantial list but don't worry, we won't change all of them.

First I'm going to adjust my *text*. I'd prefer to have my title a little larger and in Times New Roman font (which R maps to serif). I'd also like the axis labels to be in italics.

```{r}
ggplot(data = gpa_study_hours, 
       aes(x = gpa))+
  geom_histogram(color = "White", binwidth = 0.25)+
  labs(
    title = "Distribution of Grade Point Average",
    subtitle = "For 193 Introductory Statistics Students",
    x = "GPA",
    y = "Number of Students",
    caption = "Source: Private US University, 2012"
  )+
  theme(
    title = element_text(size = 16,
                              family = "serif"),
    axis.title = element_text(face = "italic")
  )
```

Next, I'm going to adjust the *panel* the plot is on. I'll change the background color and the grid. 

```{r}
ggplot(data = gpa_study_hours, 
       aes(x = gpa))+
  geom_histogram(color = "White", binwidth = 0.25)+
  labs(
    title = "Distribution of Grade Point Average",
    subtitle = "For 193 Introductory Statistics Students",
    x = "GPA",
    y = "Number of Students",
    caption = "Source: Private US University, 2012"
  )+
  theme(
    title = element_text(size = 16,
                         family = "serif"),
    axis.title = element_text(face = "italic"),
    panel.background = element_rect(fill = "lightblue"), 
    panel.grid.major.x = element_blank(), 
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = "blue"), 
    panel.grid.minor.y = element_blank()
  )
```

My next step will be to address the *plot* background (the white space behind the blue). I'd like it to be the same color. I'll also adjust the color and fill of the bars. 

```{r}
ggplot(data = gpa_study_hours, 
       aes(x = gpa))+
  geom_histogram(fill = "aquamarine4",
                 color = "lightblue", 
                 binwidth = 0.25)+
  labs(
    title = "Distribution of Grade Point Average",
    subtitle = "For 193 Introductory Statistics Students",
    x = "GPA",
    y = "Number of Students",
    caption = "Source: Private US University, 2012"
  )+
  theme(
    title = element_text(size = 16,
                         family = "serif"),
    axis.title = element_text(face = "italic"),
    panel.background = element_rect(fill = "lightblue"), 
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(color = "blue"),
    panel.grid.minor.y = element_blank(),
    plot.background = element_rect(fill = "lightblue")
  )
```