---
title: "Stat242:Lab3"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
description: >
  Building a basic linear regression model.
vignette: >
  %\VignetteIndexEntry{Stat242:Lab3}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(Stat242)
library(tidyverse)
library(moderndive)
library(openintro)
```

## Building a Linear Model

For this lab we will be trying to predict the difference between the starting price and final price (not including shipping) for a Mario Kart game auctioned on ebay. Let's take a look at our data set. 
```{r}
glimpse(mariokart)
```

We want to predict the difference of the start and final price of the game itself. Looking at the pricing variables given we have start_Pr (starting price), ship_Pr (shipping price) and total_Pr (total price). This means we need to do a bit of work to get the variable we want. Below I'll clean the data - to get rid of unnecessary variables - and add the new variable. 

```{r}
model_data <- mariokart %>% 
  select(n_bids, duration ,start_pr, ship_pr, total_pr) %>% #This line selects the variable we might need for our model.
  mutate(price_diff = total_pr-ship_pr-start_pr) #This line calculates the difference between start and final price for each auction.
```

## Finding the Correlation Coefficient

Next, we'll use a function from the `moderndive` package to calculate a few linear correlation coefficients and see if there are any relationships worth modeling. 

```{r}
get_correlation(data = model_data, price_diff ~ duration)

get_correlation(data = model_data, price_diff ~ n_bids)
```

## Looking for a Better Fit

Based on the correlation values, it would be a better choice to make our prediction based on the number of bids than the duration of the auction. While the correlation value is not spectacular, it is close enough to 1 based on the sample of 143 auctions. Before we build our model let's sketch a scatter plot. 

```{r}
ggplot(model_data, aes(n_bids, price_diff))+
  geom_point()
```

There is definitely a positive linear relationship in this data. There is also a clear outlier. Let's find that data point in the original data set. The auction title reveals that this was for a Wii with Guitar Hero 5 and Mario Kart bundled together. Let's remove this as the price of the console itself is heavily influencing the model. Once removed, we'll build a new model and compare their accuracy via correlation. 

```{r}
model_data_2 <- model_data %>% 
  filter(price_diff<300)

get_correlation(data = model_data_2, price_diff ~ n_bids)
```
This is a much better correlation. We'll use this dataset to build our model. 

## Building the Linear Model

```{r}
price_model <- lm(data = model_data_2, price_diff ~ n_bids)
```

Next we'll overlay the model onto our scatter plot. 

```{r}
ggplot(model_data_2, aes(n_bids, price_diff))+
  geom_point()+
  geom_smooth(method = "lm",
              se = FALSE)
```

## Analyzing the Model

Now we'll look at a summary of the model. 

```{r}
get_regression_table(price_model)
```

The *intercept* row is for the y-intercept of the line. Based on our data, the model shows an intercept of 12.702 and a standard error of 2.725; This leads to a confidence interval of (7.315, 18.090). For this row we aren't concerned about the (test) statistic or p_value. 

The *nBids* row is for the slope of the line. Based on our data, the model shows a slope of 1.737 (meaning the price difference goes up $1.74 per bid) and a standard error of .186; This leads to a confidence interval of (1.370, 2.104). For this row, we are concerned with the (test) statistic and p_value. The null hypothesis would be that the variables are not correlated and the slope is 0. With a test statistic of 4.662 and p_value (close to) 0, we can definitely reject the null hypothesis and support the claim that there is a relationship between these variables with a nonzero slope.

## Making Predictions

Now that our model is built, let's make some predictions. To do this, we'll use the `linear_model_prediction` function from our `Stat231` package.

We'll make two predictions: an interpolation (within the original data cluster) and an extrapolation (a point away from the original data cluster). 

Interpolation: 
```{r}
linear_model_prediction(price_model, 15)
```

Based on our model, an ebay auction for Mario Kart that receive 15 bids will have a $38.76 increase from the starting price to the final price. 

Extrapolation: 
```{r}
linear_model_prediction(price_model, 100)
```

Based on our model, an ebay auction for Mario Kart that receive 100 bids will have a $186.40 increase from the starting price to the final price. 