---
title: "Stat231:Lab6"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
description: >
  For this lab, we will be peaking at section 7.5 (without worry about all the formulas). The analysis of variance - anova - test is equivalent to the chi-squared test but for numeric data instead of categorical data; It will compare the means of three or more samples. 
vignette: >
  %\VignetteIndexEntry{Stat231:Lab6}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Stat231)
library(tidyverse)
library(openintro)
```

## ANOVA

For this lab, we will be peaking at section 7.5 (without worry about all the formulas). The analysis of variance - anova - test is equivalent to the chi-squared test but for numeric data instead of categorical data; It will compare the means of three or more samples. 

We will be going back to the diamonds data set and running an anova test to see if there is a difference in the average price of diamonds based on clarity. We will nest the test inside a *summary()* function to get the test and print out of results in one line. 


```{r}
summary(aov(price ~ clarity, data = diamonds))
```

The important parts of this print out are the $F$ value and $Pr(>F)$. The $F$ value is similar to a $z$ or $t$ score. The $Pr(>F)$ is a p-value. 

Using the p-value method, $<2e-16$ means the p-value is less than $0.0000000000000002$. Using any of the typical values for alpha, we would reject our null hypothesis and can say there is a difference in the average price of diamonds based on their clarity rating.

Using the classic method, we need to calculate a critical F value. 

```{r}
qf(0.95, #1-alpha
   df1 = 7,
   df2 = 53932)
```

This means our critical region starts at approximately $F = 2$, and our test statistic is $F = 215$. We are well into the critical region and again conclude that there is a difference in the average price of diamonds based on their clarity. 